---
title: "Signature extraction from Canadas et al 2018"
output: html_notebook
---

# Setup
```{r setup}
# install.packages("readxl")
library(readxl)
source("https://bioconductor.org/biocLite.R")
biocLite("Biobase")
biocLite("GEOquery")
biocLite("limma")
```

# Import data

```{r}
# Version info: R 3.2.3, Biobase 2.30.0, GEOquery 2.40.0, limma 3.26.8
# R scripts generated  Fri Aug 31 00:53:50 EDT 2018


# Server: www.ncbi.nlm.nih.gov
# Query: acc=GSE45120&platform=GPL6244&type=txt&groups=ZR-75-1|ZR-75-30&colors=dfeaf4|f4dfdf&selection=XXXXXX&padj=fdr&logtransform=auto&columns=ID&columns=adj.P.Val&columns=P.Value&columns=t&columns=B&columns=logFC&columns=Gene+symbol&columns=Gene+title&num=250&annot=ncbi

# Unable to generate script analyzing differential expression.
#      Invalid input: at least two groups of samples should be selected.

################################################################
#   Boxplot for selected GEO samples
library(Biobase)
library(GEOquery)

# load series and platform data from GEO

canadas_data <- paste(getwd(),"../..","datasets","canadas_2018/",sep = "/")
gset_canadas0 <- getGEO("GSE45120",destdir = canadas_data)
if (length(gset_canadas0) > 1) idx <- grep("GPL6244", attr(gset, "names")) else idx <- 1
gset_canadas <- gset_canadas0[[idx]]

# # set parameters and draw the plot
# palette(c("#dfeaf4","#f4dfdf", "#AABBCC"))
# dev.new(width=4+dim(gset)[[2]]/5, height=6)
# par(mar=c(2+round(max(nchar(sampleNames(gset)))/2),4,2,1))
# title <- paste ("GSE45120", '/', annotation(gset), " selected samples", sep ='')
# boxplot(exprs(gset), boxwex=0.7, notch=T, main=title, outline=FALSE, las=2)

```

## log2 transform
```{r}
# log2 transform
exfull <- exprs(gset_canadas)
# remove NAs
gset_canadas_trsf <- gset_canadas
qxfull <- as.numeric(quantile(exfull, c(0., 0.25, 0.5, 0.75, 0.99,1.0), na.rm=T))
LogCfill <- (qxfull[5] > 100) ||
          (qxfull[6]-qxfull[1] > 50 && qxfull[2] > 0) ||
          (qxfull[2] > 0 && qxfull[2] < 1 && qxfull[4] > 1 && qxfull[4] < 2)
# invoke this if log2 normalization isn't already in effect
if (LogCfill) { 
  print("log normalizing")
  exfull[which(exfull <= 0)] <- NaN
  exprs(gset_canadas_trsf) <- log2(exfull) 
  } else if (qxfull[6] > 100) {
  # if the qxfull[6] full is extremely high...
    print("adjusting values")
    # remove extremely high values; since these are log2 fold changes, we don't want anything insanely high
    exfull[which(exfull >= 10)] <- NA
    # also remove extremely low values
    exfull[which(exfull < -10)] <- NA
    exprs(gset_canadas_trsf) <- exfull
}

# remove all rows that are all NA's
gset_canadas_trsf_with_NA <- apply(X = gset_canadas_trsf,
                           MARGIN = 1,
                           function(x) {
                             #length(x[is.na(x)]) > 1
                             length(x[is.na(x)]) == length(x)
                             })
gset_canadas_trsf_with_NA <- gset_canadas_trsf_with_NA[gset_canadas_trsf_with_NA]

```

# Differential expression analysis 

The data came from a one-channel microarray

```{r print_description}
print(gset_canadas$description)
```

## Set design 
This is a pretty simple experimental setup: two conditions, one H69M and one H69 (both cell lines). We should consider how lung-cancer inflammation responses vary from EOC, Breast, or Colon cancer responses (as in Li et al 2014 and Chiappinelli et al 2015)

```{r set_design}
gset_canadas_meta <- data.frame(celltype=gset_canadas$`cell line:ch1`)
gset_canadas_trsf$description <- gset_canadas_meta$celltype
design_canadas <- model.matrix(~ 0 + description, gset_canadas_trsf)
colnames(design_canadas) <- levels(gset_canadas_meta$celltype)
```

## Model fit

We will now fit the model. This doesn't appear to be log2-normalized, so we will do so here
```{r}
exprs(gset_canadas_trsf) <- log2(exfull)
lmfit_canadas <- lmFit(gset_canadas_trsf, design_canadas)
```

## DEX genes

Extract necessary statistics and the differentially-expressed genes
```{r extract_stats}
alpha <- 0.05
log2fc_lim <- 0.5
canadas_cont_matrix <- makeContrasts(H69M-H69, levels=design_canadas)
lmfit_canadas_fit1 <- contrasts.fit(lmfit_canadas, canadas_cont_matrix)
lmfit_canadas_ebayes <- eBayes(fit = lmfit_canadas_fit1,proportion = alpha)
tT_canadas <- topTable(lmfit_canadas_ebayes, 
                       adjust="fdr", 
                       sort.by="B", number=nrow(gset_canadas_trsf))

# now for the frustrating business of extracting the names of different genes
tT_canadas_gene_names <- sapply(tT_canadas$gene_assignment,
                                function(x) {
                                  a <- unlist(strsplit(x = as.character(x),split = " // ",fixed = TRUE))
                                  a[2]
                                })
tT_canadas$GENE_SYMBOL <- as.character(tT_canadas_gene_names)

tT_canadas_subset <- subset(tT_canadas, select=c("ID","adj.P.Val","P.Value","t","B","logFC","GENE_SYMBOL","AveExpr"))

# # fit1 <- contrasts.fit(fit,cont.matrix)
# fit2 <- contrasts.fit(fit_bycell, cont.matrix)
# # fit1 <- eBayes(fit1, 1)
# fit2 <- eBayes(fit2, 0.01)
# tT <- topTable(fit2, adjust="fdr", sort.by="B", number=250)
# # allT <- topTable(fit2, adjust="none", sort.by="B",number = nrow(fit2))
# tT <- topTable(fit2, adjust="fdr", sort.by="B",number = nrow(fit2))

```

Volcano plot of the results
```{r volc_plot}
plot(tT_canadas_subset$logFC,
     -log10(tT_canadas_subset$adj.P.Val),
     xlab=c("log2FC (H69M-H69)"),
     ylab="adj.P.val",
     pch=20)
tT_canadas_sig_up_h69m <- subset(tT_canadas_subset,adj.P.Val <= alpha & logFC > log2fc_lim)
tT_canadas_sig_down_h69m <- subset(tT_canadas_subset,adj.P.Val <= alpha & logFC < -log2fc_lim)
points(tT_canadas_sig_up_h69m$logFC,
       -log10(tT_canadas_sig_up_h69m$adj.P.Val),
       pch=20,
       col="red")
     #col=as.factor(tT_canadas_subset$adj.P.Val <= alpha & tT_canadas_subset$logFC >= log2fc_lim))
abline(h = -log10(alpha),
       v = log2fc_lim,
       lty=2,
       col="red")

# IFNB has a fold-change in expression of 0.2
```

## Intersect the Canadas set 

Canadas et al focused on the 452 most differentially-expressed genes in their analysis (which they obtained from an earlier paper, Canadas et al Clin Cancer Res 2014). The genes are provided in Supplementary Dataset 1, though the exact method to reproduce them is not given.

```{r import_supp_dat_1}

read_excel_allsheets <- function(filename, tibble = FALSE) {
  # from https://stackoverflow.com/questions/12945687/read-all-worksheets-in-an-excel-workbook-into-an-r-list-with-data-frames 
    sheets <- excel_sheets(filename)
    x <- lapply(sheets, function(X) readxl::read_excel(filename, sheet = X))
    if(!tibble) x <- lapply(x, as.data.frame)
    names(x) <- sheets
    x
}

canadas_topgenes_sem <- read_excel_allsheets(paste(canadas_data,"41591_2018_116_MOESM3_ESM.xlsx",sep = "/"))

canadas_topgenes_sem_up <- canadas_topgenes_sem$`S2_up_down genes H69M vs H69`$`H69M-UP`
canadas_topgenes_sem_down <- canadas_topgenes_sem$`S2_up_down genes H69M vs H69`$`H69M-DOWN`

```

Replot the volcano but with the specific genes marked
```{r}
plot(tT_canadas_subset$logFC,
     -log10(tT_canadas_subset$adj.P.Val),
     xlab=c("log2FC (H69M-H69)"),
     ylab="adj.P.val",
     pch=20)
# tT_canadas_sig_up_h69m <- subset(tT_canadas_subset,adj.P.Val <= alpha & logFC > log2fc_lim)
# tT_canadas_sig_down_h69m <- subset(tT_canadas_subset,adj.P.Val <= alpha & logFC < -log2fc_lim)
points(tT_canadas_sig_up_h69m$logFC,
       -log10(tT_canadas_sig_up_h69m$adj.P.Val),
       pch=20,
       col="red")
     #col=as.factor(tT_canadas_subset$adj.P.Val <= alpha & tT_canadas_subset$logFC >= log2fc_lim))

tT_canadas_sig_up_h69m_in_sem <- subset(tT_canadas_sig_up_h69m,GENE_SYMBOL %in% canadas_topgenes_sem_up)
tT_canadas_sig_up_in_sem <- subset(tT_canadas_subset,GENE_SYMBOL %in% canadas_topgenes_sem_up)
  #subset(tT_canadas_subset,GENE_SYMBOL %in% canadas_topgenes_sem_up | GENE_SYMBOL %in% canadas_topgenes_sem_down)

points(tT_canadas_sig_up_in_sem$logFC,
       -log10(tT_canadas_sig_up_in_sem$adj.P.Val),
       pch=20,
       col="orange")


abline(h = -log10(alpha),
       v = log2fc_lim,
       lty=2,
       col="red")

```

Heatmap:

```{r hmap}

# gene annotation
gene_annot <- data.frame(sig_up_h69m = as.factor(tT_canadas_subset$GENE_SYMBOL %in% tT_canadas_sig_up_h69m$GENE_SYMBOL),
                         sig_up_h69m_and_sem = as.factor(tT_canadas_subset$GENE_SYMBOL %in% tT_canadas_sig_up_h69m_in_sem$GENE_SYMBOL),
                         sig_up_sem = as.factor(tT_canadas_subset$GENE_SYMBOL %in% tT_canadas_sig_up_in_sem$GENE_SYMBOL))

rownames(gene_annot) <- tT_canadas_subset$ID

# sample annotation
sample_annot <- gset_canadas_meta
rownames(sample_annot) <- colnames(exprs(gset_canadas_trsf))

# pare down the expression matrix to a desired subset
exp_full <- exprs(gset_canadas_trsf)
gset_canadas_trsf_hmap_sig <- exp_full[rownames(exp_full) %in% rownames(tT_canadas_sig_up_h69m),]
gene_annot_sig <- gene_annot[rownames(gene_annot) %in% rownames(gset_canadas_trsf_hmap_sig),]

# heatmap
pheatmap(gset_canadas_trsf_hmap_sig,
         annotation_col = sample_annot,
         annotation_row = gene_annot_sig,
         show_rownames = FALSE,
         show_colnames = FALSE)

# how many genes are in theirs but not mine?
# genes that are significant to them but not to me

gene_annot_up_sem_uniq <- subset(gene_annot,sig_up_sem == TRUE & sig_up_h69m == FALSE)
# match the canadas-unique probes to genes
gene_annot_up_sem_uniq_genes <- subset(tT_canadas_subset,ID %in% rownames(gene_annot_up_sem_uniq))
gene_annot_up_sem_uniq.n_genes <- unique(gene_annot_up_sem_uniq_genes$GENE_SYMBOL)
print(paste("They report",nrow(gene_annot_up_sem_uniq),"probes representing",length(gene_annot_up_sem_uniq.n_genes),"genes as significantly-upregulated in H69M that I do not"))

gene_annot_up_mine_uniq <- tT_canadas_sig_up_h69m[!tT_canadas_sig_up_h69m$ID %in% tT_canadas_sig_up_in_sem$ID,]

print(paste("I report",nrow(gene_annot_up_mine_uniq),"probes representing",length(unique(gene_annot_up_mine_uniq$GENE_SYMBOL))," genes as significantly-upregulated in H69M that they do not"))

```

Fair concordance between my analysis and theirs. 

Save my signatures
```{r}
write.table(x = tT_canadas_sig_up_in_sem,file = "./tT_canadas_sig_up_in_sem.txt",row.names = TRUE,col.names = TRUE,sep = "\t",quote = FALSE)
write.table(x = tT_canadas_sig_up_h69m,file = "./tT_canadas_sig_up_h69m.txt",row.names = TRUE,col.names = TRUE,sep = "\t",quote = FALSE)
write.table(x = tT_canadas_sig_up_h69m_in_sem,file = "./tT_canadas_sig_up_h69m_in_sem.txt",row.names = TRUE,col.names = TRUE,sep = "\t",quote = FALSE)
```



# GSVA analysis

Now, we will examine gene sets
