---
title: "Sheng and LaFleur 2018 RNA-seq reanalysis (mouse)"
output: html_notebook
---

Sheng and LaGleur used RNA-seq. They aligned their data using STAR to hg19 (it appears ENSEMBL, judging from the gene names given), yielding Integer counts. They did differential expression analysis using edgeR. 

The RNA-seq methods description from their methods

```
Raw reads were aligned to hg19 or mm9 using STAR (v2.4.2a) with the parameter ‘‘quantMode’’ set as ‘‘GeneCounts’’ to produce count table for each gene. Differential gene analysis was performed on gene raw counts in R with edgeR package (v3.18.1) from bio- conductor. Read count table was filtered so that genes with at least one count across conditions were kept. The negative binomial generalized log-linear model was used in differential analysis. A FDR cut-off of 0.05 was used to determine significantly differentially expressed genes. The R package gProfileR (v0.6.4) was used to perform gene enrichment analysis on differential genes. Gene set enrichment analysis (GSEA) was performed with R package clusterProfiler (v3.4.4).
The function analyzeRepeats.pl from Homer (http://homer.ucsd.edu/homer/) software was used to get raw counts for repeats from RNA-seq data. Differential expression for repeats was performed with edgeR the same way as for genes.
```

This notebook depicts the re-analysis of mouse data

# Setup
```{r}
install.packages("readxl")
library(readxl)
source("https://bioconductor.org/biocLite.R")
biocLite("Biobase")
biocLite("GEOquery")
biocLite("limma")
biocLite("edgeR")
biocLite("gProfileR")
biocLite("GSVA")
biocLite("GO.db")
biocLite("org.Hs.eg.db")
biocLite("org.Mm.eg.db")
source("../../pipelines/gsva_script.R")
biocLite("EnsDb.Hsapiens.v86")
```

```{r}
library("Biobase")
library("GEOquery")
library("limma")
library("edgeR")
library("gProfileR")
library("GSVA")
library("GO.db")
library("org.Hs.eg.db")
library("org.Mm.eg.db")
library("EnsDb.Hsapiens.v86")
```


# Get data

## import datasets

```{r get_csv_counts}
mouse_rnaseq <- read.csv(file = "../../datasets/sheng_lafleur/GSE112230_2018-03-01_rnaseq_mm_counts.csv",header = TRUE,row.names = 1,sep = ",")
mouse_erv <- read.csv(file = "../../datasets/sheng_lafleur/GSE112230_2018-03-22_repeat_count_mm.csv",header = TRUE,row.names = 1,sep = ",")
```

## design

Setup from GSEA
```
GSM3061399	ctl_1: Scramble rep1
GSM3061400	ctl_2: Scramble rep2
GSM3061401	ctl_4: Scramble rep3
GSM3061402	ko_5: LSD1 KO rep1
GSM3061403	ko_6: LSD1 KO rep2
GSM3061404	ko_7: LSD1 KO rep3
GSM3061405	dko_8: LSD1/MDA5 DKO rep1
GSM3061406	dko_9: LSD1/MDA5 DKO rep2
GSM3061407	dko_10: LSD1/MDA5 DKO rep3

```

```{r}
#mouse_rnaseq_design <- data.frame()
mouse_rnaseq_design <- data.frame(cell = factor(rep("B16",ncol(mouse_rnaseq))))
rownames(mouse_rnaseq_design) <- colnames(mouse_rnaseq)
mouse_rnaseq_design$treatment <- factor(c(rep("A_Scramble",3),
                                          rep("B_LSD1_KO",3),
                                          rep("C_LSD1_MDA5_DKO",3)))
mouse_rnaseq_design$rep <- factor(rep(c(1,2,3),3))
print(mouse_rnaseq_design)

# model matrix
mouse_rnaseq_model <- model.matrix(~ treatment,
                                          mouse_rnaseq_design,contrasts.arg = mouse_rnaseq_design$treatment)
print(mouse_rnaseq_model)
# treat the sh-Control as the "baseline", so it gets an intercept term
```

# edgeR analysis

## dge list
set DGElist object
```{r dgelist}
# do some filtering first
mouse_rnaseq_filt_ind <- apply(X = mouse_rnaseq,MARGIN = 1,function(X) {sum(X) > 0})
mouse_rnaseq_filt <- mouse_rnaseq[mouse_rnaseq_filt_ind,]

# upper-case the names
rownames(mouse_rnaseq_filt) <- toupper(rownames(mouse_rnaseq_filt))

# now start estimating dispersions
mouse_rnaseq_dge <- DGEList(counts=mouse_rnaseq_filt,
                            genes = data.frame(genes = rownames(mouse_rnaseq_filt))) # set DGE list
mouse_rnaseq_dge <- calcNormFactors(object = mouse_rnaseq_dge) # set normalization factors
mouse_rnaseq_dge <- estimateDisp(y = mouse_rnaseq_dge,
                                 design = mouse_rnaseq_model) # set dispersion factors
plotBCV(y = mouse_rnaseq_dge)
# mouse_rnaseq_dge_exactTest <- exactTest(mouse_rnaseq_dge)

```

## glm
fit GLM
```{r}
# The negative binomial generalized log-linear model was used in differential analysis.
mouse_rnaseq_dge_fit <- glmFit.default(y = mouse_rnaseq_dge,
                                       design = mouse_rnaseq_model,
                                       dispersion = mouse_rnaseq_dge$trended.dispersion)

```

## check conditions
Perform tests

LSD1 KO vs Ctrl
```{r}
pval_cutoff <- 0.05
n_toptag <- 10000
logFC_edgeR_lim <- 1
# LSD KO vs Ctrl

mouse_rnaseq_dge_fit_lrt_KO <- glmLRT(mouse_rnaseq_dge_fit, 
                                      coef=2)

mouse_rnaseq_dge_fit_lrt_KO_top <- topTags(mouse_rnaseq_dge_fit_lrt_KO,
                                           n = 100000,
                                           adjust.method = "BH")
mouse_rnaseq_dge_fit_lrt_KO_top_up <- mouse_rnaseq_dge_fit_lrt_KO_top
mouse_rnaseq_dge_fit_lrt_KO_top_up$table <- subset(mouse_rnaseq_dge_fit_lrt_KO_top$table,
                                                abs(logFC) > logFC_edgeR_lim & PValue <= pval_cutoff) # FDR <= pval_cutoff
# mouse_rnaseq_dge_fit_qlf_KO_top$table <- subset(mouse_rnaseq_dge_fit_qlf_KO_top$table,PValue < pval_cutoff & logFC > logFC_edgeR_lim)
 
plot(x = mouse_rnaseq_dge_fit_lrt_KO_top$table$logFC,
     y = -log10(mouse_rnaseq_dge_fit_lrt_KO_top$table$PValue),
     xlim=c(-15,15),
     pch=20,
     main=paste("Gene expression changes with LSD1 KO in mouse B16 tumors\n",
                  nrow(mouse_rnaseq_dge_fit_lrt_KO_top_up),
                  "of",
                  nrow(mouse_rnaseq_dge_fit_lrt_KO_top),
                  "genes enriched"),
     xlab=c("log2FC (log2(LSD1 KO) - log2(Scramble))"),
     ylab=c("-log10(PValue)"))

points(x = mouse_rnaseq_dge_fit_lrt_KO_top_up$table$logFC,
       y = -log10(mouse_rnaseq_dge_fit_lrt_KO_top_up$table$PValue),
       col="red",
       pch=20)
 
```


Try again with contrast matrix
```{r}
CvsKO <- c(-1,1,0)
KOvsC <- c(1,-1,0)
KOvsC_mat <- matrix(data = c(KOvsC,CvsKO),nrow = 3,ncol = 2)
mouse_rnaseq_dge_fit_lrt_KO_test <- glmLRT(mouse_rnaseq_dge_fit,contrast = KOvsC_mat)

plot(mouse_rnaseq_dge_fit_lrt_KO_test$table$logFC.1,
     -log10(mouse_rnaseq_dge_fit_lrt_KO_test$table$PValue),
     xlim=c(-40,40))
points(mouse_rnaseq_dge_fit_lrt_KO_test$table$logFC.2,
     -log10(mouse_rnaseq_dge_fit_lrt_KO_test$table$PValue),
     xlim=c(-40,40))

mouse_rnaseq_dge_fit_lrt_KO_test_top <- topTags(object = mouse_rnaseq_dge_fit_lrt_KO_test,n = 10000,
                                                adjust.method = "fdr",p.value = 0.05)

points(mouse_rnaseq_dge_fit_lrt_KO_test_top$table$logFC.1,
     -log10(mouse_rnaseq_dge_fit_lrt_KO_test_top$table$PValue),col="red")
points(mouse_rnaseq_dge_fit_lrt_KO_test_top$table$logFC.2,
     -log10(mouse_rnaseq_dge_fit_lrt_KO_test_top$table$PValue),col="blue")

```

Mirrors the images. Let's not use this...

DKO vs Ctrl
```{r}
pval_cutoff <- 0.05
n_toptag <- 10000
logFC_edgeR_lim <- 1
# LSD KO vs Ctrl

mouse_rnaseq_dge_fit_lrt_DKO <- glmLRT(mouse_rnaseq_dge_fit, 
                                      coef=3)

mouse_rnaseq_dge_fit_lrt_DKO_top <- topTags(mouse_rnaseq_dge_fit_lrt_DKO,
                                           n = 100000,
                                           adjust.method = "BH")
mouse_rnaseq_dge_fit_lrt_DKO_top_up <- mouse_rnaseq_dge_fit_lrt_DKO_top
mouse_rnaseq_dge_fit_lrt_DKO_top_up$table <- subset(mouse_rnaseq_dge_fit_lrt_DKO_top$table,
                                                abs(logFC) > logFC_edgeR_lim & PValue <= pval_cutoff) #  & FDR <= pval_cutoff

plot(x = mouse_rnaseq_dge_fit_lrt_DKO_top$table$logFC,
     y = -log10(mouse_rnaseq_dge_fit_lrt_DKO_top$table$FDR),
     xlim=c(-15,15),
     pch=20)

points(x = mouse_rnaseq_dge_fit_lrt_DKO_top_up$table$logFC,
       y = -log10(mouse_rnaseq_dge_fit_lrt_DKO_top_up$table$FDR),
       col="red",
       pch=20)
 
```

Visualize the heatmaps

```{r}
mm_dex_gene_names <- unique(c(rownames(mouse_rnaseq_dge_fit_lrt_KO_top_up$table),
                              rownames(mouse_rnaseq_dge_fit_lrt_DKO_top_up$table)))
mouse_rnaseq_filt_dex <- mouse_rnaseq_filt[rownames(mouse_rnaseq_filt) %in% toupper(mm_dex_gene_names),]
pheatmap(mat = log2(mouse_rnaseq_filt_dex+1),
         show_colnames = FALSE,
         fontsize = 5.0,
         cluster_cols = FALSE,
         annotation_col = mouse_rnaseq_design[,c(2,3)])

# show all HLA signatures
mouse_rnaseq_filt_h2 <- mouse_rnaseq_filt[grep(pattern = "H2-",x = rownames(mouse_rnaseq_filt)),]
pheatmap(mat = log2(mouse_rnaseq_filt_h2+1),
         show_colnames = FALSE,
         fontsize = 5.0,
         annotation_col = mouse_rnaseq_design[,c(2,3)])

# get the counts specific to KO-vs-Ctrl

mouse_rnaseq_design_koc <- mouse_rnaseq_design[grep(pattern = "DKO",x = mouse_rnaseq_design$treatment,invert = TRUE),]
mouse_rnaseq_filt_dex_koc <- mouse_rnaseq_filt_dex[rownames(mouse_rnaseq_filt_dex) %in% toupper(rownames(mouse_rnaseq_dge_fit_lrt_KO_top_up)),rownames(mouse_rnaseq_design_koc)]
mouse_rnaseq_design_dkoc <- mouse_rnaseq_design[grep(pattern = "_KO",x = mouse_rnaseq_design$treatment,invert = TRUE),]
mouse_rnaseq_filt_dex_dkoc <- mouse_rnaseq_filt_dex[rownames(mouse_rnaseq_filt_dex) %in% toupper(rownames(mouse_rnaseq_dge_fit_lrt_DKO_top_up)),rownames(mouse_rnaseq_design_dkoc)]

# heatmaps
pheatmap(mat = log2(mouse_rnaseq_filt_dex_koc+1),
         show_colnames = FALSE,
         fontsize = 5.0,
         cluster_cols = FALSE,
         annotation_col = mouse_rnaseq_design[,c(2,3)])

pheatmap(mat = log2(mouse_rnaseq_filt_dex_dkoc+1),
         show_colnames = FALSE,
         fontsize = 5.0,
         cluster_cols = FALSE,
         annotation_col = mouse_rnaseq_design[,c(2,3)])

```


# GO analysis

We already have the gene symbols, 
## GSVA

Some helpful code notes
```
vvv3@login04:~/pillaifolder/rmarkdown/tfh_diffn$ grep -s "gsva" *.Rmd
ranking_enhancerCluster_genes.Rmd:                             "pipelines/transcriptomics_pipeline/gsea/gsva_script.R"),
tfhdiff.Rmd:source("pipelines/transcriptomics_pipeline/gsea/gsva_script.R")
tfhdiff.Rmd:prepare_gsva()
tfhdiff.Rmd:tcell.genesets.tfh.enrich <- gsva_enrich(counts = tcell.enrich.counts$counts.aggregate,geneset = tcell.genesets)
vvv3@login04:~/pillaifolder/rmarkdown/tfh_diffn$ grep -s "tcell.genesets" *.Rmd
tfhdiff.Rmd:tcell.genesets <- import_gmt(gmtfile = "referenceAnnotation/hg38/genesets/msigdb/custom/msigdb_tcell_genesets.gmt")
tfhdiff.Rmd:tcell.genesets.tfh.enrich <- gsva_enrich(counts = tcell.enrich.counts$counts.aggregate,geneset = tcell.genesets)
tfhdiff.Rmd:#colnames(tcell.genesets.tfh.enrich$es.obs) <- c("CD4naive","PD1loCXCR5pos","PD1intCXCR5pos","PD1hiCXCR5hi")
tfhdiff.Rmd:tcell.genesets.tfh.enrich.es.kmeans <- kmeans_intervals(matrix = tcell.genesets.tfh.enrich$es.obs,kmax = 5,B=50)
tfhdiff.Rmd:tcell.genesets.tfh.enrich.es <- as.data.frame(tcell.genesets.tfh.enrich$es.obs)
tfhdiff.Rmd:tcell.genesets.tfh.enrich.es$k.choice <- tcell.genesets.tfh.enrich.es.kmeans$matrix.kmeans$kmeans
tfhdiff.Rmd:length(tcell.genesets.tfh.enrich.es$k.choice)
tfhdiff.Rmd:dim(tcell.genesets.tfh.enrich.es)
tfhdiff.Rmd:tcell.genesets.tfh.enrich.es <- tcell.genesets.tfh.enrich.es[order(tcell.genesets.tfh.enrich.es$k.choice),]
tfhdiff.Rmd:#enrichments <- tcell.genesets.tfh.enrich$es.obs[sort(tcell.genesets.tfh.enrich.es.kmeans$matrix.kmeans$kmeans),]
tfhdiff.Rmd:rowannotation.ssgsea <- data.frame(kchoice = as.factor(sort(tcell.genesets.tfh.enrich.es.kmeans$matrix.kmeans$kmeans)))
tfhdiff.Rmd:#rowannotation.ssgsea <- data.frame(kchoice = as.factor(tcell.genesets.tfh.enrich.es.kmeans$matrix.kmeans$kmeans))
tfhdiff.Rmd:head(tcell.genesets.tfh.enrich.es)
tfhdiff.Rmd:pheatmap(mat = tcell.genesets.tfh.enrich.es[,c(1:(ncol(tcell.genesets.tfh.enrich.es)-1))],
vvv3@login04:~/pillaifolder/rmarkdown/tfh_diffn$
```

```{r}
# prepare for GSVA
prepare_gsva()
```

```{r}
# import mouse gene sets
mm_c2_genesets <- import_gmt(gmtfile = "../../referenceAnnot/genesets/mm10/msigdb_c2_mmusculus.gmt")
mm_c6_genesets <- import_gmt(gmtfile = "../../referenceAnnot/genesets/mm10/msigdb_c6_mmusculus.gmt")
mm_c7_genesets <- import_gmt(gmtfile = "../../referenceAnnot/genesets/mm10/msigdb_c7_mmusculus.gmt")

# capitalize the gene sets
rownames(mouse_rnaseq_filt) <- toupper(rownames(mouse_rnaseq_filt))
rownames(mouse_rnaseq_filt_dex) <- toupper(rownames(mouse_rnaseq_filt_dex))

# perform GSVA with GO terms from Li et al Oncotarget
mm_c2_genesets_enrich <- gsva_enrich(counts = mouse_rnaseq_filt_dex_koc,
                                 geneset = mm_c2_genesets)
mm_c6_genesets_enrich <- gsva_enrich(counts = mouse_rnaseq_filt_dex_koc,
                                 geneset = mm_c6_genesets)
mm_c7_genesets_enrich <- gsva_enrich(counts = mouse_rnaseq_filt_dex_koc,
                                 geneset = mm_c7_genesets)
mm_c5_bp_enrich <- gsva_enrich(counts = mouse_rnaseq_filt_dex_dkoc,
                               geneset = bp_genesets)

# perform GSVA with Li oncotarget signatures
mm_interferon_enrich <- gsva_enrich(counts = mouse_rnaseq_filt_dex_koc,
                                 geneset = interferon_genesets)
mm_cytochemokine_enrich <- gsva_enrich(counts = mouse_rnaseq_filt_dex_koc,
                                 geneset = cytochemokine_genesets)

# perform GSVA with viral terms
mm_virus_enrich <- gsva_enrich(counts = mouse_rnaseq_filt_dex_koc,
                                 geneset = virus_genesets)

# perform GSVA with hallmark terms
mm_hallmark_enrich <- gsva_enrich(counts = mouse_rnaseq_filt_dex_koc,
                                 geneset = hallmark_genesets)



```


Set the terms of the signatures that we want
```{r}
# set li oncotarget terms
li_oncotarget_terms <- paste("interferon",
                             "cytokine",
                             "inflammation",
                             "influenza",
                             "testes",
                             "presentation",
                             sep = "|")
```

Plot enrichment
```{r}
pheatmap(mm_c2_genesets_enrich,
         annotation_col = mouse_rnaseq_design,
         show_colnames = FALSE,
         cluster_cols = FALSE,
         fontsize = 5.0,
         main = "C2")
pheatmap(mm_c6_genesets_enrich,
         annotation_col = mouse_rnaseq_design,
         show_colnames = FALSE,
         cluster_cols = FALSE,
         fontsize = 5.0,
         main = "C6")
pheatmap(mm_c7_genesets_enrich,
         annotation_col = mouse_rnaseq_design,
         show_colnames = FALSE,
         show_rownames = FALSE,
         fontsize = 5.0,
         main = "C7")

```

Not much enrichment. 

```{r}
# highlight the immunological GO_BP terms
mm_c5_bp_li_oncotarget <- grep(pattern = li_oncotarget_terms,x = rownames(mm_c5_bp_enrich),ignore.case = TRUE)
mm_c5_bp_not_li_oncotarget <- grep(pattern = li_oncotarget_terms,x = rownames(mm_c5_bp_enrich),ignore.case = TRUE,invert = TRUE)
mm_c5_bp_annot <- data.frame(li_oncotarget = c(rep("li_onco",length(mm_c5_bp_li_oncotarget)),
                                            rep("non_li_onco",length(mm_c5_bp_not_li_oncotarget))))
rownames(mm_c5_bp_annot) <- c(rownames(mm_c5_bp_enrich)[mm_c5_bp_li_oncotarget],
                           rownames(mm_c5_bp_enrich)[mm_c5_bp_not_li_oncotarget])
#c5_bp_annot <- c5_bp_annot[order(c5_bp_annot$),]
mm_c5_bp_sorted <- mm_c5_bp_enrich[rownames(mm_c5_bp_annot),]
pheatmap(mm_c5_bp_sorted,
         annotation_col = mouse_rnaseq_design,
         annotation_row = mm_c5_bp_annot,
         show_rownames = FALSE,
         show_colnames = FALSE,
         fontsize = 5.0,
         main = "BP")

```


The Li Oncotarget signatures in mouse
```{r}
pheatmap(mm_interferon_enrich,
         annotation_col = mouse_rnaseq_design,
         labels_row = rownames(mm_interferon_enrich),
         show_colnames = FALSE,
         cluster_cols = TRUE,
         fontsize = 2.5,
         main = "Interferon GO Terms")
pheatmap(mm_cytochemokine_enrich,
         annotation_col = mouse_rnaseq_design,
         labels_row = rownames(mm_cytochemokine_enrich),
         show_colnames = FALSE,
         fontsize = 5.0,
         cluster_cols = FALSE,
         main = "Cytokine/Chemokine GO Terms")
pheatmap(mm_virus_enrich,
         annotation_col = mouse_rnaseq_design,
         labels_row = rownames(mm_virus_enrich),
         show_rownames = TRUE,
         show_colnames = FALSE,
         cluster_cols = FALSE,
         fontsize_row = 2.5,
         main = "Virus GO Terms")
pheatmap(mm_hallmark_enrich,
         annotation_col = mouse_rnaseq_design,
         labels_row = rownames(mm_hallmark_enrich),
         show_rownames = TRUE,
         show_colnames = FALSE,
         cluster_cols = FALSE,
         fontsize_row = 2.5,
         main = "Hallmark GO Terms")



```

Not a whole lot of condition-specific enrichment in other pathways.

